Cultural Transmission and Technological Transitions during the Late Paleolithic in Korea

### 2022-08-25 analysis/paper/paper-doc.qmd

PCA serves to examine correlation of attributes for determining the dominant mode of cultural transmission. @fig-five shows positive relationship between attributes. Especially, there are strong relationships between body length and maximum length, tang width and mid width, and tang length and stem length. Body length and maximum length contribute the most to the shape of stemmed points and they are most correlated to each other. Other width and tang related attributes are close to each other but a bit far from body length and maximum length. Stem width contributes the least to the overall shape. We can assume that the body length and maximum length were prioritized when the stemmed points were transmitted.

![PCA analysis for stemmed points. The direction of the arrow shows the relationship between attributes. The length and color of the arrow indicate the strength of contribution to the stemmed point shape.](images/007-PCA-1.png){#fig-five width="284"}

4.  Variation Over Time

@@

6.  Variation Between Raw Materials

>>>>

PCA serves to examine correlation of attributes for determining the dominant mode of cultural transmission. @fig-pca-attributes shows positive relationship between attributes. Especially, there are strong relationships between body length and maximum length, tang width and mid width, and tang length and stem length. Body length and maximum length contribute the most to the shape of stemmed points and they are most correlated to each other. Other width and tang related attributes are close to each other but a bit far from body length and maximum length. Stem width contributes the least to the overall shape. We can assume that the body length and maximum length were prioritized when the stemmed points were transmitted.

```{r}
#| label: fig-pca-attributes
#| fig-cap: PCA analysis for stemmed points, for all phases combined, then for only phase two and phase three. The direction of the arrow shows the relationship between attributes; arrows pointing in the same direction indicate variables that are highly postitively correlated. The length of the arrow indicates the strength of contribution to the stemmed point shape.
#| fig-width:   7
#| fig-height:  3

library(ggbiplot)
library(ggfortify)
### PCA analysis using ggbiplot, ggfortify pkgs

PCA <- prcomp(DF, center = TRUE, scale = TRUE)

# remove stage 1 artefact
xx_pca <-
  df_sitename %>%
  filter(SPstage.Stage != 1) %>%
  select(-sitename, -SPstage.Stage, -SPstage.Raw_material) %>%
  prcomp(., center = TRUE, scale = TRUE)

xx_pca1 <-
  data.frame(xx_pca$x,
             stage = as.factor( df_sitename$SPstage.Stage[df_sitename$SPstage.Stage !=1]))

## CV with site (n=X)
label_phase <-
  df_sitename %>%
  group_by(SPstage.Stage) %>%
  add_tally() %>%
  mutate(label = as.factor(paste0(SPstage.Stage,'\nN = ',n)))

count_phase <- length(which(label_phase[9]>0))
count_phase2 <- length(which(label_phase[9] == 2))
count_phase3 <- length(which(label_phase[9] == 3))

###

# phase 2

xx_pca2 <-
  df_sitename %>%
  filter(SPstage.Stage == 2) %>%
  select(-sitename, -SPstage.Stage, -SPstage.Raw_material) %>%
  prcomp(., center = TRUE, scale = TRUE)

# phase 3

xx_pca3 <-
  df_sitename %>%
  filter(SPstage.Stage == 3) %>%
  select(-sitename, -SPstage.Stage, -SPstage.Raw_material) %>%
  prcomp(., center = TRUE, scale = TRUE)

##### using factoextra pkg
library(factoextra) #http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/112-pca-principal-component-analysis-essentials/

# Color variables by the continuous variable
cv_all <-
fviz_pca_var(xx_pca) +
  ggtitle(paste0("All phases", " (n = ", count_phase,")")) +
  theme(legend.position = "none")

# phase 2
cv_two <-
  fviz_pca_var(xx_pca2) +
  ggtitle(paste0("Phase 2", " (n = ", count_phase2,")"))

# phase 3
cv_three <-
  fviz_pca_var(xx_pca3) +
  ggtitle(paste0("Phase 3", " (n = ", count_phase3,")"))

library(patchwork)
cv_all + cv_two + cv_three
```

## Variation Over Time

@@

##  Variation Between Raw Materials
